---
title: "Final Analyses"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

# Data Import

### Libraries

```{r Packages, results='hide', message=FALSE}
library(tidyverse)
library(lme4)
library(nlme)
library(psych)
library(finalfit)
library(naniar)
library(recipes)
library(stringr)
library(reshape2)
library(sjPlot)
library(emmeans)
library(purrr)
library(plotly)
library(jtools)
library(knitr)
library(kableExtra)
library(survey)

here::here()
```

```{r Plotting Themes}
set_theme(base = theme_minimal())
```


### Load Data

```{r Load Data}
load("data.rdata")
load("personality_measures.rData")
```

### Column Names

```{r fix colnames, echo = FALSE}
nlsy <- rename(nlsy,
       extrav = extra_tipi,
       agree = agree_tipi,
       consc = consc_tipi,
       open= open_tipi,
       neuro = neuro_tipi,
       race = race1
)
nlsy_l <- rename(nlsy_l,
       extrav = extra_tipi,
       agree = agree_tipi,
       consc = consc_tipi,
       open= open_tipi,
       neuro = neuro_tipi,
       race = race1
)
nlsy <- nlsy %>% 
  select(pid:neuro_mi, race)

nlsy <- nlsy %>% 
  mutate(sex = tolower(sex) %>% 
           factor(),
         race1 = race,
         race = forcats::fct_collapse(race_full,
                                        other = c("", "aapi", 
                                                  "indig", "other")))

nlsy_l <- nlsy_l %>% 
  select(pid:neuro_mi, race:school_group)

add <- add %>% 
  mutate(ever_suspended = str_to_title(ever_suspended) %>% 
           factor(),
    pid = as.character(pid) %>% 
      as.numeric()
  )
nlsy$race <- nlsy$race %>% 
  forcats::fct_relevel("black" , "white" , "latinx", "other")
  
```

### Sampling Weights

```{r Sampling Weights}
#################### SAMPLE WEIGHTS  
###### NLSY
nlsy_w <- rio::import(here::here("data/nlsy_weights/customweight_nlscya_6260ed2f1804a9e8208.dat")) %>% 
  as_tibble()
names(nlsy_w) <- c("pid", "weight")
nlsy <- left_join(nlsy, nlsy_w, by = "pid") %>% 
  mutate(
    pid = factor(pid)
  )

###### NLSY LONG
nlsy_l <- full_join(nlsy_l, nlsy_w, by = "pid") %>% 
  mutate(
    pid = factor(pid)
  )


#str(nl)
###### ADD
add_w <- rio::import(here::here("data/add_weights.sav"))
names(add_w) <- c("pid", "cluster", "weight_long", "weight", "weight_long_w134")
add_w <- add_w %>% 
  mutate(
    pid = as.character(pid) %>% 
      as.numeric()
  )
add <- left_join(add, add_w, by = "pid") %>% 
  mutate(
    pid = factor(pid)
  )
```

### Final Variable Counts: 

NLSY: 18
ADD Health: 23

```{r manipulating Data into lists}
#str(add)
data <- list(nlsy, add)

# CREATING MAP-ABLE LIST OF DATA

n <- nlsy %>% melt(
  id.vars = c("pid", "sex", "birth_year", "race", "race_full", "race1", "age_tipi", "ever_suspended", "weight"),
  variable.name = "measure", 
  value.name = "score") %>% 
  split(., .$measure)

a <- add %>% melt(
  id.vars = c("pid", "sex", "birth_year", "race", "hispanic", "home_language", 
              "US_born", "grade_w1", "suspension_this_year", "ever_suspended",
              "grade_last_suspended", "pid", "cluster", "weight_long", "weight", "weight_long_w134"),
  variable.name = "measure", 
  value.name = "score") %>% 
  split(., .$measure)



d <- c(a, n)
names_list <- names(d)
names_data <- c(rep("_add", length(a)), rep("_nlsy", length(n)))
names(d) <- paste0(names_list, names_data)
```

### lists for Sampling Weights Adjusted

```{r}
require(survey)
nlsy_adj <- svydesign(id     = ~pid,
                     weights = ~weight,
                     data    = nlsy)

nlsy_l_adj <- svydesign(id   = ~pid,
                     weights = ~weight,
                     data    = nlsy_l)

add_adj <- svydesign(id      = ~pid,
                     strata  = ~cluster,
                     weights = ~weight,
                     data    = add)
data_adj <- list(nlsy_adj, add_adj)
names(data_adj) <- c("nlsy_adj", "add_adj")
```

```{r manipulating weighted data into lists}
# CREATING MAP-ABLE LIST OF DATA
a_adj <- map(a, ~svydesign(id     = ~pid,
                           weights = ~weight,
                           strata  = ~cluster,
                           data    = .x))

n_adj <- map(n, ~svydesign(id     = ~pid,
                           weights = ~weight,
                           data    = .x))

d_adj <- c(a_adj, n_adj)
names_list <- names(d_adj)
names_data_adj <- c(rep("_add_adj", length(a_adj)), rep("_nlsy_adj", length(n_adj)))
names(d_adj) <- paste0(names_list, names_data)
```

# Descriptive Statistics

## Demographics (Sampling)

```{r Participant Demographics Table}
############### Pretty Table Function
pretty_table <- function(df){
  x <- deparse(substitute(df))
  df %>% 
  kbl(caption = x) %>%
  kable_classic(full_width = F, html_font = "Cambria")
}

############### FUNCTION
pdem <- function(df){
  df %>% 
  group_by(race, sex) %>% 
  count(.) %>% 
  pretty_table(.)
}

################ ANALYSIS
pdem(nlsy)
pdem(add)
```

```{r Participant Demographics Figure}
############### FUNCTION
pdem_plot <- function(df){
  df %>% 
  ggplot() +
  geom_bar(aes(str_to_title(race), fill = str_to_title(sex)),
           position = "dodge") +
  theme_apa() +
  theme(legend.position = "bottom") +
  labs(
    x = "Race",
    y = "N"
  )
}
################ ANALYSIS
pdem_plot(nlsy)
pdem_plot(add)
```

## Proportion Suspended (Unweighted)

```{r}
################ FUNCTION
susp_prop <- function(df){
  proportion <- df %>% 
  group_by(sex, race, ever_suspended) %>% 
  count() %>% 
  na.omit()
proportion <- spread(proportion, key = ever_suspended, value = n) %>% 
  tibble(.)
proportion$total <- proportion$No + proportion$Yes
proportion$prop <- round(proportion$Yes/proportion$total, 2)
colnames(proportion) <- c("Sex", "Race", "Not Suspended", "Suspended", 
                          "Total n", "Proportion")
return(proportion)
}


# PROPORTION PLOT

plot_prop <- function(df){
  df %>% 
    mutate(Demographic = paste(str_to_title(df$Race), str_to_title(df$Sex), sep = "\n"),
           Percent = Proportion * 100) %>% 
  ggplot() +
    geom_col(aes(Demographic, Percent, fill = Race),
             position = "dodge") +
  labs(title = "Proportion of Students Suspended",
       caption = deparse(substitute(df)),
       subtitle = "By Race and Gender",
       x = "Race",
       y = "Percent of students") +
    ylim(0, 75) +
  theme_apa() +
  theme(legend.position = "none")
}


###### NLSY
nlsy_prop <- susp_prop(nlsy)
pretty_table(nlsy_prop)
plot_prop(nlsy_prop)


####### ADD HEALTH
add_prop <- susp_prop(add)
pretty_table(add_prop)
plot_prop(add_prop)
```

## (W) Proportion Suspended 

```{r}
require(survey)
########### Creating Data Table
adj_samples <- map(data_adj, ~svytotal(
  ~interaction(.x$variables$sex, .x$variables$race, .x$variables$ever_suspended),
  .x
))

########### 
## NLSY HEALTH
nlsy_pt <- tibble(
total_no = c(51745256, 56610495, 143520099, 104834469, 17886824, 19258129, 580087226, 549189747), 
total_yes = c(5180151, 17243098, 50680215, 86714845, 2654794, 5961711, 35086329, 94804143),
SE_no = c(3423172, 3872039, 4689369, 4114781, 1698220, 1718261, 13940592, 13800227),
SE_yes = c(1150771, 2049397, 3026543, 3941767, 58024, 1091982, 3574711, 5876934),
race = c(  "other",  "other", 
           "black",  "black",
           "latinx", "latinx", 
           "white",  "white") %>% 
  factor(),

sex = c("female", "male", "female", "male", 
        "female", "male", "female", "male") %>% 
  factor()
)
nlsy_pt <- nlsy_pt %>% 
  mutate(
    Percent =total_yes/(total_yes + total_no))
  

########### 
## ADD HEALTH

add_pt <- tibble(
total_no = c(720038, 1076668, 5324124, 6522058, 731884, 972598, 263620, 332552),
total_yes = c(1000375, 681941, 2417160, 975055, 476266, 246398, 113382, 49626), 
SE_no = c(54634, 57848, 156621, 151210, 60623, 68309, 32481, 38782),
SE_yes = c(68223, 50460, 117643, 70422, 50586, 33480, 29711, 14093),
race = c("black", "black", 
         "white", "white", 
         "latinx", "latinx", 
         "other", "other") %>% 
  factor(),

sex = c("male", "female", "male", "female",  
        "male", "female", "male", "female") %>% 
  factor()
)
add_pt <- add_pt %>% 
  mutate(
    Percent = total_yes/(total_yes + total_no))

######## LIST
adj_prop <- list(nlsy_pt, add_pt)
names(adj_prop) <- c("NLSY_adj", "ADD_adj")

########## Output
map(adj_prop, pretty_table)
########## Adj Plot Function
adj_plot_f <- function(x)
  {x %>% 
    ggplot() +
    geom_col(aes(race, Percent, fill = sex),
             position = "dodge") +
  labs(title = "Proportion of Students Suspended",
       caption = deparse(substitute(x)),
       subtitle = "By Race and Gender",
       x = "Race",
       y = "Percent of students") +
    ylim(0, 1) +
  theme_apa() +
  theme(legend.position = "none")
}

map(adj_prop, adj_plot_f)
```


# Inferential Statistics

## Race x Gender inxn Effects Of Being Suspended 

```{r Race x Gender Interaction Function, results='asis'}
race_by_gender_interaction <- function(df){
  name <- deparse(substitute(df))
  rsm <- glm(ever_suspended ~ race * sex,
             family = binomial, data = df)
  sum <- summary(rsm)
  plot <- plot_model(rsm, type = "pred", terms = c("race", "sex"))
  emm <- emmeans(rsm, specs = pairwise ~ race | sex,
                 type = "response", adjust = "none")
  assign(
    x = paste0(name, "_rxg"),
    value = list(rsm, sum, plot, emm),
    envir = .GlobalEnv
  )
}

race_by_gender_interaction(nlsy)
race_by_gender_interaction(add)
```

> This is a reproduction of prior analytic work that shows yes, there are *large racial differences in suspension rate*.I believe the ADD-Health data are, however, a novel "replication"

## (W) Race x Gender inxn Effects Of Being Suspended 

```{r}
race_by_gender_interaction_adj <- function(df){
  name <- deparse(substitute(df))
  rsm <- svyglm(ever_suspended ~ race * sex,
                family = quasibinomial, 
                design = df)
  sum <- summary(rsm)
  plot <- plot_model(rsm, 
                     type = "pred", 
                     terms = c("race", "sex"))
  emm <- emmeans(rsm, specs = pairwise ~ race | sex,
                 type = "response", adjust = "none")
  assign(
    x = paste0(name, "_rxg"),
    value = list(rsm, sum, plot, emm),
    envir = .GlobalEnv
  )
}

race_by_gender_interaction_adj(nlsy_adj)
race_by_gender_interaction_adj(add_adj)
```

## Output for race x gender x suspension models

```{r}
print(nlsy_rxg)
print(nlsy_adj_rxg)
print(add_rxg)
print(add_adj_rxg)
```

# Analysis: 

## Big Five Moderation Analysis

```{r Personality models and outputs}
# CREADING LISTS OF OUR VARIOUS OUTPUTS
p_mod <- map(d, ~glm(ever_suspended ~ race * sex * score,
             family = binomial, 
             data = .x))
p_sum <- map(p_mod, summary)
p_plot <- map(p_mod, ~plot_model(.x, type = "pred", 
                                 terms = c("score", "race", "sex"),
                                 colors = c("red1", "deepskyblue2", "gray 79", "grey 66")))
p_emtrend_output <- map(p_mod, ~emtrends(.x, pairwise ~ race*sex, 
                                         var = "score", 
                                         adjust = "none"))
##### Merging lists
pers_models <- list(NULL)
for(i in seq_along(p_mod)){
  pers_models[[i]] <-  list(p_mod[[i]], p_sum[[i]], p_plot[[i]], p_emtrend_output[[i]])
}
# Setting sub-component names
for(i in seq_along(pers_models)){
  names(pers_models[[i]]) = c("Model", "Summary", "Plot", "Emtrend_Output")
}
# Setting names
names(pers_models) = paste0(names_list, names_data)
```

```{r Searching for just p<.05}


sig_values_funct(p_emtrend_output)

sig_values <- list(NULL)
for(i in seq_along(p_emtrend_output)){
  sig_values[[i]] = summary(p_emtrend_output[[i]]$contrasts) %>% 
    as_tibble(.) %>% 
    filter(`p.value` < .05)}

names(sig_values) = paste0(names_list, names_data)
```

## (W) Big Five Moderation Analysis

```{r Personality models and outputs}
# CREADING LISTS OF OUR VARIOUS OUTPUTS
p_mod_adj <- map(d_adj, ~svyglm(ever_suspended ~ race * sex * score,
             family = binomial, 
             design = .x))
            
w <- plot_model(p_mod_adj$extrav_add, type = "pred", 
                terms = c("score", "race", "sex"))
 
p_sum_adj <- map(p_mod_adj, summary)
p_plot_adj <- map(p_mod_adj, ~plot_model(.x, type = "pred", 
                                 terms = c("score", "race", "sex"),
                                 colors = c("red1", "deepskyblue2", "gray 66", "grey 79")))

for(i in 1:18){
  p_mod_adj[i][["naive.cov"]] <- NULL
  p_mod_adj[i][["cov.unscaled"]] <- NULL
  p_mod_adj[i][["survey.design"]] <- NULL
}

emtrends(p_mod_adj$extrav_add, pairwise ~ race*sex,
         var = "score", 
         adjust = "none",
         weights = p_mod_adj$extrav_add$model$`(weights)`,
         data = p_mod_adj$extrav_add$data)


p_emtrend_output_adj <- map(p_mod_adj, ~emtrends(.x, 
                                                 pairwise ~ race*sex,
                                                 var = "score", 
                                                 adjust = "none",
                                                 weights = .x$model$`(weights)`,
                                                 data = .x$data)
)
##### Merging lists
pers_models_adj <- list(NULL)
for(i in seq_along(p_mod_adj)){
  pers_models_adj[[i]] <-  list(p_mod_adj[[i]], p_sum_adj[[i]], p_plot_adj[[i]], p_emtrend_output_adj[[i]])
}
# Setting sub-component names
for(i in seq_along(pers_models_adj)){
  names(pers_models_adj[[i]]) = c("Model", "Summary", "Plot", "Emtrend_Output")
}
`# Setting names
names(pers_models_adj) = paste0(names_list, names_data_adj)

# Setting sub-component names
for(i in seq_along(pers_models_adj)){
  names(pers_models_adj[[i]]) = c("Model (weighted)", "Summary (weighted)", "Plot #(weighted)", "Emtrend_Output (weighted)")
}
# Setting names
names(pers_models_adj) = paste0(names_list, names_data_adj)
```

```{r Searching for just p<.05}
sig_values_adj <- list(NULL)
for(i in seq_along(p_emtrend_output_adj)){
  sig_values_adj[[i]] = summary(p_emtrend_output_adj[[i]]$contrasts) %>% 
    as_tibble(.) %>% 
    filter(`p.value` < .05)}

names(sig_values_adj) = paste0(names_list, names_data_adj)

```


# Sensitivity Analysis: Examining Variance with Loaded Models

```{r Variance Loaded Modeling Function}
var_loaded_funct <- function(model, trait){
  plot <<- plot_model(model, 
           type = "pred", 
           terms = c(paste0(trait), "race", "sex"))
  output <<- emtrends(model, 
                    pairwise ~ race*sex, 
                    var = paste0(trait), 
                    adjust = "none")
}
```


## NLSY TIPI Variance-Loaded Models

```{r NLSY TIPI Variance Models, eval = F}
# Model 
nlsy_tipi_variance_model_adj <- svyglm(ever_suspended ~ 
                                  race * sex * extrav + 
                                  race * sex * open + 
                                  race * sex * neuro + 
                                  race * sex * consc + 
                                  race * sex * agree,
             family = binomial, design = nlsy_adj)

# summary(nlsy_tipi_variance_model)

nlsy_tipi_loaded_adj <- list(NULL)
trait <- c("extrav", "agree", "neuro", "consc", "open")
for(i in seq_along(trait)){
  var_loaded_funct(nlsy_tipi_variance_model_adj, paste0(trait[i]))
  nlsy_tipi_loaded_adj[[paste0(trait[i])]][[1]] <- plot
  nlsy_tipi_loaded_adj[[paste0(trait[i])]][[2]] <- output
}

nlsy_tipi_loaded

```

## NLSY MINI-IPIP Variance-Loaded Models

```{r NLSY MINI-IPIP Variance Models, eval = F}
## Model 
nlsy_mini_ipip_variance_model_adj <- svyglm(ever_suspended ~ 
                                  race * sex * extrav_mi + 
                                  race * sex * agree_mi + 
                                  race * sex * neuro_mi + 
                                  race * sex * consc_mi + 
                                  race * sex * open_mi,
                                  family = binomial, 
                                  design = nlsy_adj)

# summary(nlsy_mini_ipip_variance_model)

nlsy_mini_ipip_loaded <- list(NULL)
trait <- c("extrav_mi", "agree_mi", "neuro_mi", "consc_mi", "open_mi")
for(i in seq_along(trait)){
  var_loaded_funct(nlsy_mini_ipip_variance_model, paste0(trait[i]))
  nlsy_mini_ipip_loaded[[paste0(trait[i])]][[1]] <- plot
  nlsy_mini_ipip_loaded[[paste0(trait[i])]][[2]] <- output
}

nlsy_mini_ipip_loaded
```

## ADD Health Variance-Loaded Models

```{r Add Health Variance Models, eval = F}
# Model 
add_variance_model_adj <- svyglm(ever_suspended ~ 
                            race * sex * open + 
                            race * sex * extrav + 
                            race * sex * agree + 
                            race * sex * neuro + 
                            race * sex * consc,
                            family = binomial, 
                            design = add_adj)

summary(add_variance_model)

add_variance_loaded <- list(NULL)
trait <- c("extrav", "agree", "neuro", "consc", "open")
for(i in seq_along(trait)){
  var_loaded_funct(add_variance_model, paste0(trait[i]))
  add_variance_loaded[[paste0(trait[i])]][[1]] <- plot
  add_variance_loaded[[paste0(trait[i])]][[2]] <- output
}

add_variance_loaded
```

## Explanatory Power

```{r}
#install and load pscl package
library(pscl)

#calculate McFadden's R-squared for model

pR2(nlsy_tipi_variance_model_adj)['McFadden']
```


# Longitudinal Analyses

```{r Longitudinal Models and Output }
nl <- nlsy_l %>% 
  select(pid:neuro_mi, race2:weight)%>% 
  melt(
  id.vars = c("pid", "sex", "birth_year", "race", "race_full", "race2", "age_tipi", "ever_suspended", "weight", "school_group", "years", "value", "school_age"),
  variable.name = "measure", 
  value.name = "score") %>% 
  split(., .$measure)

# CREADING LISTS OF OUR VARIOUS OUTPUTS
l_mod <- map(nl, ~glm(value ~ race * sex * score * school_group,
             family = binomial, data = .x))
l_sum <- map(l_mod, summary)
l_plot <- map(l_mod, ~plot_model(.x, type = "pred", 
                                 terms = c("score", "race", "sex", "school_group")))
l_trends <- map(l_mod, ~emtrends(.x, pairwise ~ race * sex * school_group, 
                     var = "score", adjust = "none")
)
##### Merging lists
l_models <- list(NULL)
for(i in seq_along(l_mod)){
  l_models[[i]] <-  list(l_mod[[i]], 
                         l_sum[[i]], 
                         l_plot[[i]], 
                         l_trends[[i]])
}
# Setting sub-component names
for(i in seq_along(l_models)){
  names(l_models[[i]]) = c("Model", "Summary", "Plot", "Emtrend_Output")
}
# Setting names
names(l_models) <- names(nl)
```

```{r}
sig_values <- list(NULL)
for(i in seq_along(l_trends)){
  sig_values[[i]] = summary(l_trends[[i]]$contrasts) %>% 
    as_tibble(.) %>% 
    filter(`p.value` < .05)
}
names(sig_values) = names(nl)
print(sig_values)
```

# Analysis for presentations

```{r}
sigval_ordered <- sig_values[order(names(sig_values))]
sigval_ordered_adj <- sig_values_adj[order(names(sig_values_adj))]
mods_ordered <- pers_models[order(names(pers_models))]
plots_ordered <- p_plot[order(names(p_plot))]
plots_ordered_adj <- p_plot_adj[order(names(p_plot_adj))]
```

```{r Race X Gender Bar Chart}
n <- nlsy_adj_rxg[[3]]$data  %>% 
  tibble()
n <- n %>% 
  mutate(
         race = factor(n$x, 
                       labels = c("Black", "White", "Latinx", "Other")),
         dataset = "NLSY")
a <- add_adj_rxg[[3]]$data  %>% 
  tibble()
a <- a %>% 
  mutate(
         race = factor(a$x, labels = c("Black", "White", "Latinx", "Other")),
         dataset = "ADD Health")
d <- rbind(n, a)
d$group <- d$group %>% 
  forcats::fct_relevel("male", "female")
levels(d$group) <- c("Male", "Female")
d %>%                  
  ggplot(aes(x = group, predicted, fill = race)) +
  geom_col(position = "dodge") +
  geom_errorbar(aes(x = group, ymin = conf.low, ymax = conf.high),
                position = position_dodge(width = .9),
                width = .2,
                size = .5,
                color = "gray35") +
  scale_fill_manual(values = c("red1", "deepskyblue2", "gray 66", "grey 79")) +
  labs(fill = "Race") +
  facet_wrap(~dataset) +
  ylab("Suspension Rate") +
  xlab("Gender") +
  
  theme_minimal() 

```


```{r}
display_funct <- function(i){
  sigval_ordered[paste0(i)] %>% 
    print
  plots_ordered[paste0(i)] %>% 
    print
  plots_ordered_adj[paste0(i)] %>% 
    print
}
display_funct("optimism_add")
```

```{r}
p_only <- svyglm(ever_suspended ~ open + extrav + agree + neuro + consc + anger + anxiety + optimism,
           family = binomial, 
           design = add_adj)
p_only <- svyglm(ever_suspended ~ open + extrav + agree + neuro + consc,
           family = binomial, 
           design = nlsy_adj)

pR2(p_only)['McFadden']

```


