---
title: "Final Analyses"
author: "Zach Schroeder"
date: "10/12/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = FALSE, warning = FALSE)
options(digits = 3)
```

```{r Packages, results='hide', message=FALSE}
library(tidyverse)
library(lme4)
library(nlme)
library(psych)
library(finalfit)
library(naniar)
library(recipes)
library(stringr)
library(reshape2)
library(sjPlot)
library(emmeans)
source("cleaning_functions.R")

here::here()
```

# Data Import and Cleaning

## NLSY Data Preparation 

### Import and Pre-written cleaning

```{r NLSY Import, results='hide', message=FALSE}
new_data <- rio::import(here::here("data", "NLSY_11_19_21/NLSY_11_19_21.csv"))
```

> NLSY is cleaned using NLS Investigator Code included with data download from NLSY site. The code is 1600 lines long and not included in the markdown output. Please see the .Rmd file for NLSY cleaning code

```{r NLSY Cleaning, echo=FALSE, cache=TRUE}
names(new_data) <- names_from_cleaning_functions

# Handle missing values

  new_data[new_data == -1] = NA  # Refused 
  new_data[new_data == -2] = NA  # Dont know 
  new_data[new_data == -3] = NA  # Invalid missing 
  new_data[new_data == -7] = NA  # Missing 


# If there are values not categorized they will be represented as NA

pid <- new_data$C0000100
new_data <- vallabels(new_data)
new_data <- qnames(new_data)
attr(new_data , "variable.labels") <- varlabels
nlsy_full <- new_data
rm(new_data)

nlsy_full$pid <- pid
```

### Race/Ethnicity

#### Simplified Race Variable

```{r NLSY Creating race variable from provided simple race information}
# 
nrace1 <- nlsy_full %>% 
  select(pid, race) %>% 
  mutate(race = factor(race) %>% 
           recode_factor(HISPANIC = "latinx",
                         BLACK = "black",
                         "NON-BLACK, NON-HISPANIC" = "other") %>% 
           forcats::fct_relevel(c("other", "black", "latinx"))
  )
names(nrace1) <- c("pid", "race1")
#table(nrace1$race)
```

#### Complex Race Variable

```{r nlsy race data import and cleaning}
### NLSY Race: They LIED to me and I feel betrayed.

new_data <- read.table(here::here('data/race_identity/race_identity.dat'), sep=' ')

### Handle missing values

  new_data[new_data == -1] = NA  # Refused 
  new_data[new_data == -2] = NA  # Dont know 
  new_data[new_data == -3] = NA  # Invalid missing 
  new_data[new_data == -7] = NA  # Missing 

names(new_data) <- race_new_data_names 


```

```{r Implementing above cleaning functions}
nr <- vallabels_race(new_data)
names(nr) <- nr_names
nr <- nr %>% 
  janitor::clean_names()
```

```{r}
nlsy_race <- nr[,c(8:38, 40:74)] # Removing unnecessary columns
#
colnames(nlsy_race) <- as.character(1:66)

nlsy_race <- nlsy_race %>% 
  unite("all", "1":"66", na.rm = T)

ncr <- gsub('_', '', nlsy_race$all) 

ncr <- gsub('NA', '', ncr) %>% 
  tolower()

#I've not got these racial markers. We should discuss what to do with them.
n <- fct_collapse(ncr,
                   white = c("white", "whiteother", "whitewhite"),
                   black = c("blackblack", "black", "blackasian", "blacklatinx",
                             "blackother", "blackapi", "blackindig", 
                             "blackindigother", "whiteblack", "whiteblackindig", "whiteblackindigasian", 
                             "whiteblackother", "blackindiglatinx", "whiteblackasian","whiteblackindigother"),
                   latinx = c("latinx", "latinxlatinx", "otherlatinx", "whitelatinx", 
                              "whiteotherlatinx", "apilatinx", "whiteindiglatinx" ),
                   aapi = c("asian", "asianother", "api", "apiother", "whiteasian", 
                            "whiteapi","whiteapiother", "asianapi"),
                   indig = c("indig","indigother", "whiteindig", "whiteindigother"),
                   other = c("other", "otherother", "whiteasianindig")
)

### 94-98 were a different scale

r1994 <- factor(new_data$Y0009800, 
    levels=c(1.0,2.0,3.0,4.0,5.0,6.0), 
    labels=c("black",
      "white",
      "latinx",
      "indig",
      "aapi",
      "other"))
r1996 <- factor(new_data$Y0396800, 
    levels=c(1.0,2.0,3.0,4.0,5.0,6.0), 
    labels=c("black",
      "white",
      "latinx",
      "indig",
      "aapi",
      "other"))
r1998 <- factor(new_data$Y0686600, 
    levels=c(1.0,2.0,3.0,4.0,5.0,6.0), 
    labels=c("white",
      "black",
      "indig",
      "Asian",
      "aapi",
      "other"))
  
r <- tibble(n, r1994, r1996, r1998)
r <- r %>% 
  unite("all", n:r1998)
r$all <- gsub('NA', '', r$all)
r$all <- gsub('_', '', r$all) 
r$all <- gsub(' ', '', r$all)
x <- factor(r$all) %>% 
  fct_collapse(.,
                   white = c("white", "whitewhiteother", "otherwhite", 
                             "whiteother", "whitewhite", "whitewhitewhite"),
                   black = c("black", "blackblack", "blackother", "blackwhite", 
                             "indigblack", "otherblack", "latinxblack", "whiteblack", 
                             "whitewhiteblack", "aapiblack"),
                   latinx = c("latinx", "latinxlatinx", "otherlatinx", "latinxindig", 
                              "whitelatinx", "indiglatinx", "otherlatinx", 
                              "latinxwhite", "latinxother"),
                   aapi = c("aapi","aapiaapi","aapiother", "otherAsian", "aapiwhite", "Asian"),
                   indig = c("indig", "otherindig", "whiteindig", "indigindig", "indigwhite"),
                   other = c("other", "otherother")) %>% 
  as.character(.) %>% 
  gsub(' ', '', .) %>% 
  factor()
nlsy_race <- tibble(new_data[,1], x)
colnames(nlsy_race) <- c("pid", "race_full")
nlsy_full <- left_join(nlsy_full, nlsy_race)
```


### first filter

> We're removing all participants born after 1996, because they don't have the full 18 years of data for schooling, as the final suspension year information was collectd is 2014

```{r Filtering <18 year olds}
# Removing participants who aren't 18 at most recent data collection level
nlsy <- nlsy_full %>%
   filter(!as.integer(nlsy_full$year) %in% c(18:41)) #factor levels for birth year > 1994
#nlsy$pid <- 1:nrow(nlsy)
``` 

### Personality Measures

```{r NLSY TIPI, cache=TRUE, warning = FALSE}
# Selecting TIPI scores
nlsy_loop <- nlsy %>% 
  select('Q16-9~000001_2006':'Q16-9~000010_2008', 
         "Q16-9~000001_2010":"Q16-9~000010_2010", 
         "Q16-9~000001_2012":"Q16-9~000010_2012", 
         "Q16-9~000001_2014":"Q16-9~000010_2014",  
         "Q16-9~000001_2016":"Q16-9~000010_2018")

# Creating empty matrix to load TIPI scores
first_nlsy_tipi <- matrix(NA, nrow(nlsy), 10) %>% 
  as_tibble()
year_tipi <- rep(NA, nrow(first_nlsy_tipi))

# Loop to find first full TIPI score
tipi_sections <- seq(from = 1, to = 70, by = 10)

for (j in 1:nrow(nlsy_loop)) {
  for (i in tipi_sections) {
  if (complete.cases(nlsy_loop[j,i:(i+9)]) == FALSE) { 
    next
    } else {
            first_nlsy_tipi[j, 1:10] <- nlsy_loop[j,i:(i+9)]
            year_tipi[j] <- i
            break
            }
        }
}

# Loop to create numeric variables
tipi <- matrix(NA, nrow(first_nlsy_tipi), ncol(first_nlsy_tipi))
for (i in (1:nrow(first_nlsy_tipi))) {
  for (j in (1:ncol(first_nlsy_tipi))) {
    tipi[i,j] <- str_sub(first_nlsy_tipi[i,j], start = 1, end = 1) %>% 
      as.numeric()
  }
}
# NOTE: The above loop creates warnings for str_sub() on empty cells. It is fine :-)

tipi <- as_tibble(tipi)
colnames(tipi) <- c("extra1", "agree1", "consc1", "neuro1", "open1",
                         "extra2", "agree2", "consc2", "neuro2", "open2")

tipi_items <- c("extra1", "agree1", "consc1", "neuro1", "open1",
                         "extra2", "agree2", "consc2", "neuro2", "open2")
tipi_questions <- c("EXTRAVERTED, ENTHUSIASTIC",
"CRITICAL, QUARRLESOME",
"DEPENDABLE, SELF-DISCIPLINED",
"ANXIOUS, EASILY UPSET",
"OPEN TO NEW EXPERIENCES, COMPLEX",
"RESERVED, QUIET",
"SYMPATHETIC, WARM",
"DISORGANIZED, CARELESS",
"CALM, EMOTIONALLY STABLE",
"CONVENTIONAL, UNCREATIVE")
NLSY_TIPI_QUESTIONS <- tibble(tipi_items, tipi_questions)

# Creating reverse coded keys 
keys <- c(1, -1, 1, -1, 1, -1, 1, -1, 1, -1)
tipi_r <- reverse.code(keys, tipi) %>% 
  as_tibble()

nlsy_tipi_items <- tipi_r %>% 
  select(starts_with("e"), starts_with("a"), starts_with("c"), starts_with("n"), starts_with("o"))

tipi_rev <- tipi_r %>% 
  select(ends_with("-"), everything()) %>% 
  cor(nlsy_tipi_items, use = "pairwise") %>% 
  round(2)

NLYS_TIPI_CORRELATIONS <- cor(nlsy_tipi_items, use = "pairwise") %>% 
  round(2) %>%
  knitr::kable()

tipi_r$extra_tipi <- (tipi_r$extra1 + tipi_r$`extra2-`)/2
tipi_r$agree_tipi <- (tipi_r$`agree1-` + tipi_r$agree2)/2
tipi_r$consc_tipi <- (tipi_r$consc1 + tipi_r$`consc2-`)/2
tipi_r$neuro_tipi <- (tipi_r$`neuro1-` + tipi_r$neuro2)/2
tipi_r$open_tipi <- (tipi_r$open1 + tipi_r$`open2-`)/2
tipi_r$pid <- nlsy$pid
nlsy <- left_join(nlsy, tipi_r, by = "pid")

```

```{r NLSY TIPI Correlations}
get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
  }
  
upper_tri <- get_upper_tri(tipi_rev)
upper_tri

library(reshape2)
melted_cormat <- melt(upper_tri)
library(ggplot2)
ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile()

```

```{r NLSY Mini-IPIP}
mini_ipip <- nlsy %>% 
  select(`Q16-9A~000001_2012`:`Q16-9A~000020_2012`) %>% 
  as_tibble()

# FUNCTION FOR RECODING TO NUMERIC
mini_ipip_code <- function(data){
  factor(data, levels = c("Very Inaccurate","Moderately Inaccurate",
             "Neither Accurate Nor Inaccurate", 
             "Moderately Accurate","Very Accurate"),
  labels = c("1","2","3","4","5")) %>% 
    as.numeric(data)
}

cols <- names(mini_ipip)
mini_ipip[cols] <- lapply(mini_ipip[cols], mini_ipip_code)

psych::describe(mini_ipip)

# Key for reverse coding
keys <- c(1,1,1,-1,1,
          -1,-1,-1,1,-1,
          1,1,1,-1,-1,
           -1,-1,-1,1,-1)

# REVERSE CODING
mini_ipip_r <- psych::reverse.code(keys, mini_ipip) %>% 
  as.matrix(ncol = 32) %>% 
  as_tibble()

# keeping names the same (without '-' coerced in reverse.code())
names(mini_ipip_r) <- c("EXTRAV1", "AGREE1", "CONSC1", "NEURO1", "OPEN1", 
                        "EXTRAV2", "AGREE2", "CONSC2", "NEURO2", "OPEN2",
                        "EXTRAV3", "AGREE3", "CONSC3", "NEURO3", "OPEN3",
                        "EXTRAV4", "AGREE4", "CONSC4", "NEURO4", "OPEN4"
                        )

extrav_mi = mini_ipip_r %>% 
  select(starts_with("EXTRAV")) %>% 
  rowMeans(na.rm=T)

agree_mi = mini_ipip_r %>% 
  select(starts_with("AGREE")) %>%  
  rowMeans(na.rm=T)

consc_mi = mini_ipip_r %>% 
  select(starts_with("CONSC")) %>%  
  rowMeans(na.rm=T)

neuro_mi = mini_ipip_r %>% 
  select(starts_with("NEURO")) %>% 
  rowMeans(na.rm=T)

open_mi = mini_ipip_r %>% 
  select(starts_with("OPEN")) %>% 
  rowMeans(na.rm=T)
pid = nlsy$pid

mip_nlsy <- tibble(pid, extrav_mi, open_mi, consc_mi, neuro_mi, agree_mi)

#### Table for reviewing item-level issues:
nlsy_mini_ipip_items <- mini_ipip_r %>% 
  select(starts_with("EXTRAV"), starts_with("AGREE"), starts_with("CONSC"), starts_with("NEURO"), starts_with("OPEN"))

NLYS_MINI_IPIP_CORRELATIONS <- cor(nlsy_mini_ipip_items, use = "pairwise") %>% 
  round(., 2)

nlsy_mini_ipip_items$pid <- nlsy$pid

nlsy_mini_ipip <- full_join(mip_nlsy, nlsy_mini_ipip_items, by = "pid")

psych::describe(nlsy_mini_ipip_items)



items <- c("EXTRAV1", "AGREE1", "CONSC1", "NEURO1", "OPEN1", 
                        "EXTRAV2", "AGREE2", "CONSC2", "NEURO2", "OPEN2",
                        "EXTRAV3", "AGREE3", "CONSC3", "NEURO3", "OPEN3",
                        "EXTRAV4", "AGREE4", "CONSC4", "NEURO4", "OPEN4"
                        )
questions <- c("I AM THE LIFE OF THE PARTY","I SYMPATHIZE WITH OTHERS' FEELINGS","I GET CHORES DONE RIGHT AWAY","I HAVE FREQUENT MOOD SWINGS","I HAVE A VIVID IMAGINATION","I DON'T TALK A LOT","I AM NOT INTERESTED IN OTHER PEOPLE'S PROBLEMS","I OFTEN FORGET TO PUT THINGS BACK IN THEIR PROPER PLACE","I AM RELAXED MOST OF THE TIME","I AM NOT INTERESTED IN ABSTRACT IDEAS","I TALK TO A LOT OF DIFFERENT PEOPLE AT PARTIES","I FEEL OTHERS' EMOTIONS","I LIKE ORDER","I GET UPSET EASILY","I HAVE DIFFICULTY UNDERSTANDING ABSTRACT IDEAS","I KEEP IN THE BACKGROUND","I AM NOT REALLY INTERESTED IN OTHERS","I MAKE A MESS OF THINGS","I SELDOM FEEL BLUE","I DO NOT HAVE A GOOD IMAGINATION)")

NLSY_MINI_IPIP_QUESTIONS <- tibble(items, questions)
```

Mini-IPIP Items:
I AM THE LIFE OF THE PARTY
I SYMPATHIZE WITH OTHERS' FEELINGS
I GET CHORES DONE RIGHT AWAY
I HAVE FREQUENT MOOD SWINGS
I HAVE A VIVID IMAGINATION
I DON'T TALK A LOT
I AM NOT INTERESTED IN OTHER PEOPLE'S PROBLEMS
I OFTEN FORGET TO PUT THINGS BACK IN THEIR PROPER PLACE
I AM RELAXED MOST OF THE TIME
I AM NOT INTERESTED IN ABSTRACT IDEAS
I TALK TO A LOT OF DIFFERENT PEOPLE AT PARTIES
I FEEL OTHERS' EMOTIONS
I LIKE ORDER
I GET UPSET EASILY
I HAVE DIFFICULTY UNDERSTANDING ABSTRACT IDEAS
I KEEP IN THE BACKGROUND
I AM NOT REALLY INTERESTED IN OTHERS
I MAKE A MESS OF THINGS
I SELDOM FEEL BLUE
I DO NOT HAVE A GOOD IMAGINATION


```{r NLSY TIPI & MINI-IPIP correlations}
nlsy_personality <- left_join(tipi_r, nlsy_mini_ipip, by = "pid") %>% 
  select("pid", sort(colnames(.)))
NLSY_PERSONALITY_CORMAT_FULL <- cor(nlsy_personality, use = "pairwise") %>% 
  round(2)
NLSY_PERSONALITY_COR <- nlsy_personality %>% 
  select(agree_mi, agree_tipi, extra_tipi, extrav_mi, open_mi, open_tipi, consc_mi, consc_tipi, neuro_mi, neuro_tipi) %>% 
  cor(., use = "pairwise") %>% 
  round(2)
```

> Capitalized variables are Mini-IPIP, lowercase are TIPI
We're seeing the same problems here, the reverse-coded items are massively impacting the correlations between scales. Composite score correlations range from .20 (agreeableness) to 0.45 (extraversion)

### Age

```{r NLSY Age}
# Creating variable for year when participants completed their first full TIPI
nlsy$year_tipi <- (year_tipi - 1) / 5 + 2006

nlsy <- nlsy %>% 
  mutate(
    birth_year = as.character(year)
  )

nlsy$birth_year <- str_sub(nlsy$birth_year, start = 1, end = 4) %>% 
  as.numeric()
nlsy$age_tipi <- nlsy$year_tipi - nlsy$birth_year
nlsy <- subset(nlsy, select = -c(month, year))
```

### Suspension

```{r NLSY Suspension}
nlsy_ever1 <-  nlsy[ , grepl( "ever" , names( nlsy ) ) ]
nlsy_ever <- unlist(nlsy_ever1) %>% 
  tolower()
nlsy_ever <- gsub("no", 0, nlsy_ever)
nlsy_ever <- gsub("yes", 1, nlsy_ever)
nlsy_ever <- nlsy_ever %>% 
  as.numeric() %>% 
  matrix(ncol = 14) %>% 
  as_tibble()
nlsy_ever[is.na(nlsy_ever)] <- 0
ever_suspended <- apply(nlsy_ever, 1, max, na.rm = T) %>% 
  as.numeric()
  
nlsy$ever_suspended <- factor(ever_suspended,
                              levels = c(0,1),
                              labels = c("No", "Yes"))
```

### Joining Data Frames

```{r NLSY Joining into Data Frame}
nlsy <- left_join(nlsy, nlsy_mini_ipip, by = "pid")
nlsy <- left_join(nlsy, nrace1, by = "pid") 
nlsy <- nlsy %>% 
  select(pid, race1, race_full, sex, birth_year, age_tipi, ever_suspended, extra_tipi, agree_tipi, consc_tipi, open_tipi, neuro_tipi, extrav_mi, agree_mi, consc_mi, open_mi, neuro_mi, ever_1988:grade_2014)
```

### Final Glimpse

```{r}
finalfit::ff_glimpse(nlsy)
```

## ADD Health Data Preparation

### Import and basic cleaning

```{r ADD Import, cache = TRUE, results='hide', message=FALSE}
# Import each of the sub-files in the ADD Health dataset
add_1 <- rio::import(here::here("data/ICPSR_21600/DS0001/21600-0001-Data.rda"))
add_5 <- rio::import(here::here("data/ICPSR_21600/DS0005/21600-0005-Data.rda"))
add_8 <- rio::import(here::here("data/ICPSR_21600/DS0008/21600-0008-Data.rda"))
add_16 <- rio::import(here::here("data/ICPSR_21600/DS0016/21600-0016-Data.rda"))
add_22 <- rio::import(here::here("data/ICPSR_21600/DS0022/21600-0022-Data.rda"))

#COMBINING VARIABLES
add_full <- full_join(add_1, add_5, by = "AID")
add_full <- full_join(add_full, add_8, by = "AID")
add_full <- full_join(add_full, add_16, by = "AID")
add_full <- full_join(add_full, add_22, by = "AID")

add_full <- sjlabelled::remove_all_labels(add_full)
```

```{r ADD Colnames and Factor Labels}
# CREATING NEW DATA FRAME OF JUST VARIABLES OF INTEREST 
add <- select(add_full, 
              #DEMOGRAPHICS
              AID, BIO_SEX, H1GI1M, H1GI1Y, H1GI9, 
              H1GI4, H1GI8, H1GI9, 
              H1GI10, H1GI11, H1GI20,
              # SUSPENSION
              H2ED3, H1ED7, H1ED8, 
              # PERSONALITY
              H4PE1, H4PE2, H4PE3, H4PE4, H4PE5, H4PE6, H4PE7, 
              H4PE8, H4PE9, H4PE10, H4PE11, H4PE12, H4PE13, 
              H4PE14, H4PE15, H4PE16, H4PE17, H4PE18, H4PE19, 
              H4PE20, H4PE21, H4PE22, H4PE23, H4PE24, H4PE25, H4PE26, 
              H4PE27, H4PE28, H4PE29, H4PE30, H4PE31, H4PE32,
              # ACADEMIC PERFORMANCE
              H1ED11, H1ED12, H1ED13, H1ED14,
              H1ED5, H1ED9, H1ED10, EREXITST,
              # DELINQUENCY SCALE
              H1DS1, H1DS2, H1DS3, H1DS4,
              H1DS5, H1DS6, H1DS7, H1DS8, 
              H1DS9, H1DS10, H1DS11, H1DS12,
              H1DS13, H1DS14, H1DS15,
              # INTERVIEWER NOTES
              H1IR1, H1IR2, H1IR4, H1IR5,
              # SCHOOL COVARIATES
              H1ED15, H1ED16, H1ED17, H1ED18,
              H1ED19, H1ED20, H1ED21, H1ED22, 
              H1ED23, H1ED24,
              # COLLEGE ASPIRATIONS
              H1EE1, H1EE2,
              # SCHOOL BELONGING
              H1PR2, H1PR4) 

add_names <- c("pid", "sex", "month", "birth_year", "race", 
           "hispanic", "race_if_select_more_than_one", "interviewer_coded_race", 
           "home_language", "US_born", "grade_w1", 
           "suspension_this_year", "ever_suspended", "grade_last_suspended",
           "EXTRAV1", "AGREE1", "CONSC1", "NEURO1", 
           "OPEN1", "ANXIETY1", "OPTIMISM1", "ANGER1", 
           "EXTRAV2", "AGREE2", "CONSC2", "NEURO2", 
           "OPEN2", "ANXIETY2", "OPTIMISM2", "ANGER2",
           "EXTRAV3", "AGREE3", "CONSC3", "NEURO3", 
           "OPEN3", "ANXIETY3", "OPTIMISM3", "ANGER3",
           "EXTRAV4", "AGREE4", "CONSC4", "NEURO4", 
           "OPEN4", "ANXIETY4", "OPTIMISM4", "ANGER4",
           # ACADEMIC PERFORMANCE
           "grade_english", "grade_math", "grade_history", "grade_science",
           "repeated_grade", "expelled", "grade_when_expelled",
           "hs_exit_status",
           # DELINQUENCY SCALE
           "graffiti", "damage_property", "lie_to_parents", "shoplift",
           "serious_fight", "serious_fight_doctor", "run_away", "carjack", 
           "theft_more_50dollars", "burgle", "mug_with_weapon", "deal_drugs", 
           "steal_less_50", "gang_fight", "rowdy_in_public",
           # INTERVIEWER REMARKS
           "attractive_physical", "attractive_personality", "candidness", "safe_in_r_home",
           "getting_along_with_teacher", "paying_attention", "doing_homework", "getting_along_with_students",
           "feel_close_to_people", "am_part_of_school", "students_are_prejudiced", "happy_to_be_at_school", 
           "teachers_treat_students_fairly", "feel_safe_at_school",
           # COLLEGE ASPIRATIONS
           "want_go_college", "will_go_college",
           # SCHOOL BELONGING
           "teachers_care_about_me", "friends_care_about_me")
names(add) <- add_names

# Creating subgroups for cleaning
add_mini_ipip <- add %>% 
  select(EXTRAV1:ANGER4)

add_demographics <- add %>% 
  mutate(., 
         sex = factor(sex, 
                          levels = c("(1) (1) Male",
                                     "(2) (2) Female"), 
                          labels = c("male", "female")),
         race = factor(race,
                       levels = c("(1) (1) White",
                                  "(2) (2) Black/African American",
                                  "(3) (3) American Indian/Native American",
                                  "(4) (4) Asian/Pacific Islander",
                                  "(5) (5) Other" ),
                       labels = c("White", "Black", "Indigenous/Native American", 
                                  "Asian/Pacific Islander", "Other")),
         hispanic = factor(hispanic,
                           levels = c("(0) (0) No",
                                      "(1) (1) Yes"),
                           labels = c("no", "yes")),
         home_language = factor(home_language,
                                levels = c("(1) (1) English",
                                           "(2) (2) Spanish",
                                           "(3) (3) Other"),
                                labels = c("English", "Spanish", "Other")),
         US_born = factor(US_born,
                          levels = c("(0) (0) No",
                                     "(1) (1) Yes (skip to Q.15)"),
                          labels = c("no", "yes")),
         grade_w1 = factor(grade_w1, 
                           levels = c("(07) (7) 7th grade",
                                      "(08) (8) 8th grade",
                                      "(09) (9) 9th grade",
                                      "(10) (10) 10th grade",
                                      "(11) (11) 11th grade",
                                      "(12) (12) 12th grade"),
                           labels = c(7, 8, 9, 10, 11, 12)),
         suspension_this_year = factor(suspension_this_year,
                          levels = c("(0) (0) No",
                                     "(1) (1) Yes"),
                          labels = c("no", "yes")),
         ever_suspended = factor(ever_suspended,
                          levels = c("(0) (0) No (skip to Q.9)",
                                     "(1) (1) Yes"),
                          labels = c("no", "yes")),
         grade_last_suspended = factor(grade_last_suspended,
                                       levels = c("(01) (1) 1st", 
                                                  "(02) (2) 2nd",
                                                  "(03) (3) 3rd",
                                                  "(04) (4) 4th", 
                                                  "(05) (5) 5th",
                                                  "(06) (6) 6th",
                                                  "(07) (7) 7th",
                                                  "(08) (8) 8th",
                                                  "(09) (9) 9th", 
                                                  "(10) (10) 10th",
                                                  "(11) (11) 11th", 
                                                  "(12) (12) 12th"),
                                       labels = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12))
         )

# REPLACING CHARACTERS MONTH/YEAR
add_demographics$birth_year <- str_sub(add_demographics$birth_year, start = 2, end = 3) %>% 
      as.numeric()
# REORDERING COLUMNS
add <- add_demographics %>% 
  select(pid, sex, birth_year, race, hispanic, home_language, US_born, 
         grade_w1, suspension_this_year, ever_suspended, grade_last_suspended)
``` 

### Personality (Mini-IPIP)

I AM THE LIFE OF THE PARTY
I SYMPATHIZE WITH OTHERS' FEELINGS
I GET CHORES DONE RIGHT AWAY
I HAVE FREQUENT MOOD SWINGS
I HAVE A VIVID IMAGINATION
 I worry about things
 I'm always optimistic about my future
 I get angry easily
I DON'T TALK A LOT
I AM NOT INTERESTED IN OTHER PEOPLE'S PROBLEMS
I OFTEN FORGET TO PUT THINGS BACK IN THEIR PROPER PLACE
I AM RELAXED MOST OF THE TIME
I AM NOT INTERESTED IN ABSTRACT IDEAS
 I am not easily bothered by things
 I hardly ever expect things to go my way
 I rarely get irritated
I TALK TO A LOT OF DIFFERENT PEOPLE AT PARTIES
I FEEL OTHERS' EMOTIONS
I LIKE ORDER
I GET UPSET EASILY
I HAVE DIFFICULTY UNDERSTANDING ABSTRACT IDEAS
I get stressed out easily
Overall, I expect more good things to happen to me than bad
I lose my temper
I KEEP IN THE BACKGROUND
I AM NOT REALLY INTERESTED IN OTHERS
I MAKE A MESS OF THINGS
I SELDOM FEEL BLUE
I DO NOT HAVE A GOOD IMAGINATION
I don't worry about things that have already happened
I rarely count on good things happening to me
I keep my cool

```{r ADD Mini-IPIP, warning = FALSE}
for (i in 1:ncol(add_mini_ipip)) {
    add_mini_ipip[,i] <- str_sub(add_mini_ipip[,i], start = 2, end = 2) %>% 
      as.numeric()
}

cols <- names(add_mini_ipip)

keys <- c(1,1,1,-1,1,-1,1,-1,
          -1,-1,-1,1,-1,1,-1,1,
          1,1,1,-1,-1,-1,1,-1,
           -1,-1,-1,1,-1,1,-1,1)

########## KEYS SOLVED

keys <- keys*-1

# REVERSE CODING
mini_ipip_r <- psych::reverse.code(keys, add_mini_ipip) %>% 
  as.matrix(ncol = 32) %>% 
  as_tibble()

AMICOR_Rev <- mini_ipip_r %>% 
  select(ends_with("-"), everything())%>% 
  cor(., use = "pairwise") %>% 
  round(., 2)

ADD_MINI_IPIP_COR <- mini_ipip_r %>% 
  select(starts_with("EXTRA"), starts_with("AGREE"), starts_with("CONSC"), starts_with("NEURO"), starts_with("OPEN"), starts_with("ANX"), starts_with("OPTIM"), starts_with("ANGER")) %>% 
  cor(., use = "pairwise") %>% 
  round(., 2)
# keeping names the same (without '-' coerced in reverse.code())
names(mini_ipip_r) <- cols

ami_items <- c("I AM THE LIFE OF THE PARTY","I SYMPATHIZE WITH OTHERS' FEELINGS","I GET CHORES DONE RIGHT AWAY","I HAVE FREQUENT MOOD SWINGS","I HAVE A VIVID IMAGINATION"," I worry about things"," I'm always optimistic about my future"," I get angry easily","I DON'T TALK A LOT","I AM NOT INTERESTED IN OTHER PEOPLE'S PROBLEMS","I OFTEN FORGET TO PUT THINGS BACK IN THEIR PROPER PLACE","I AM RELAXED MOST OF THE TIME","I AM NOT INTERESTED IN ABSTRACT IDEAS"," I am not easily bothered by things"," I hardly ever expect things to go my way"," I rarely get irritated","I TALK TO A LOT OF DIFFERENT PEOPLE AT PARTIES","I FEEL OTHERS' EMOTIONS","I LIKE ORDER","I GET UPSET EASILY","I HAVE DIFFICULTY UNDERSTANDING ABSTRACT IDEAS","I get stressed out easily","Overall, I expect more good things to happen to me than bad","I lose my temper","I KEEP IN THE BACKGROUND","I AM NOT REALLY INTERESTED IN OTHERS","I MAKE A MESS OF THINGS","I SELDOM FEEL BLUE","I DO NOT HAVE A GOOD IMAGINATION","I don't worry about things that have already happened","I rarely count on good things happening to me","I keep my cool")

add_mini_ipip_questions <- tibble(cols, ami_items, keys)

# CREATING SUM MINI-IPIP SCORES PER SCORING INSTRUCTIONS
extrav <- mini_ipip_r %>% 
  select(starts_with("EXTRAV")) %>% 
  rowMeans(na.rm=T)

agree = mini_ipip_r %>% 
  select(starts_with("AGREE")) %>%  
  rowMeans(na.rm=T)

consc = mini_ipip_r %>% 
  select(starts_with("CONSC")) %>%  
  rowMeans(na.rm=T)

neuro = mini_ipip_r %>% 
  select(starts_with("NEURO")) %>% 
  rowMeans(na.rm=T)

open = mini_ipip_r %>% 
  select(starts_with("OPEN")) %>% 
  rowMeans(na.rm=T)

anger = mini_ipip_r %>% 
  select(starts_with("ANGER")) %>%  
  rowMeans(na.rm=T)

optimism = mini_ipip_r %>% 
  select(starts_with("OPTIM")) %>% 
  rowMeans(na.rm=T)

anxiety = mini_ipip_r %>% 
  select(starts_with("ANXIE")) %>% 
  rowMeans(na.rm=T)

pid = add$pid

mip_clean <- tibble(pid, extrav, agree, consc, neuro, open, anger, optimism, anxiety)

ADD_MINI_IPIP_SCORES <- tibble(extrav, agree, consc, neuro, open, anger, optimism, anxiety)
```

> Solved the Keys problem! Their coding for Likert scales was reverse what they said it was.


```{r ADD MINI-IPIP Correlations}
mini_ipip_r <- mini_ipip_r %>% 
  select(starts_with("ex"), starts_with("ag"), starts_with("co"), starts_with("ne"), starts_with("ope"), starts_with("anx"), starts_with("opt"), starts_with("ang"))

ADD_MINI_IPIP_CORS <- cor(mini_ipip_r, use = "pairwise") %>% 
  round(., 2) %>% 
  psych::mat.sort()

get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
  }
  
upper_tri <- get_upper_tri(AMICOR_Rev)
upper_tri

library(reshape2)
melted_cormat <- melt(upper_tri)
library(ggplot2)
ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile()

```

### Cleaning Function 

```{r ADD cleaning, warning=FALSE}
add <- left_join(add, mip_clean, by = "pid")

# Save complete dataframes

add_cleaned <- add 

# Use function to filter cases that are incomplete for our analyses

add <- completeFun(add, c("race", "sex", "birth_year", "extrav", "agree", "neuro", "consc", "open", "ever_suspended"))
```

### Race/Ethnicity

```{r ADD Race}
# THIS GOT MESSY BECAUSE OF SOME REDUNDANCY IN THE LEVELLING OF FACTORS

race_simple <- rep(NA, nrow(add))

for (i in 1:nrow(add)) {
  if (add$race[i] == "Black"){
    race_simple[i] <- "BLACK" 
  }else {
      race_simple[i] <- "NON-BLACK, NON-HISPANIC"
    }
}

race <- paste0(add$race, add$hispanic)  %>% 
  factor()
levels(race)
race <- forcats::fct_collapse(race,
                      black = c("Blackno", "BlackNA"),
                      latinx = c("Asian/Pacific Islanderyes","Indigenous/Native Americanyes", "Otheryes", "Whiteyes", "Blackyes"),
                      white = c("WhiteNA", "Whiteno"),
                      other = c("Asian/Pacific IslanderNA","Asian/Pacific Islanderno", "Indigenous/Native Americanno" ,"NANA" ,                         "NAno","OtherNA","Otherno"))

add$race<- race

add$race <- add$race %>% 
  forcats::fct_relevel(c("black", "white", "latinx", "other"))
```

### Final Glimpse

```{r}
finalfit::ff_glimpse(add)
```

## Final Cleaning 

### Cleaning Global Environment

```{r Cleaning Global Environment}
rm(add_1, add_16, add_22, add_5, add_8, add_cleaned, add_demographics, add_full, first_nlsy_tipi, mini_ipip, mini_ipip_r, mip_clean, mip_nlsy, nlsy_ever, nlsy_full, nlsy_long2, nlsy_loop, tipi, tipi_r, upper_tri)
```

### Complete Cases

```{r Complete Cases Function applied to data}
nlsy <- completeFun(nlsy, c("race_full", "sex", "extra_tipi", "agree_tipi", "consc_tipi", "neuro_tipi", "open_tipi", "age_tipi", "ever_suspended"))

add <- completeFun(add, c("race", "sex", "birth_year", "extrav", "agree", "neuro", "consc", "open", "ever_suspended")) 

```

```{r final race variables}
nlsy$race2 <- nlsy$race_full %>% 
  fct_collapse(
    other = c("", "aapi", "indig", "other")
  ) %>% 
  forcats::fct_relevel(c("black", "white", "latinx", "other"))
levels(nlsy$race2)
add$race %>% 
  levels()
```

# Adding in additional variables 4/28/2022

## ADD Health

```{r}
# DS1
ds1_grades <- c(
  "H1ED11", # ENGLISH
  "H1ED12", # MATH
  "H1ED13", # HISTORY
  "H1ED14", # SCIENCE
  "S10A", # ENGLISH
  "S10B", # MATH
  "S10C", # HISTORY
  "S10D", # SCICNCE
)
# DS5
ds5_grades <- c(
  "H2ED7",  # ENGLISH
  "H2ED8", # MATH
  "H2ED9", # HISTORY
  "H2ED10", # SCIENCE
)

add_diploma <- c(
  "H3ED3", #HAS RECEIVED DIPLOMA
  "H3ED9",  #HAS NOT RECEIVED DIPLOMA
  "H4ED1" # Wave 4 hs graduation status
)

add_delq <- c(
  "H2DS1",
  "H2DS2",
  "H2DS3",
  "H2DS4",
  "H2DS5",
  "H2DS6",
  "H2DS7",
  "H2DS8",
  "H2DS9",
  "H2DS10",
  "H2DS11",
  "H2DS12",
  "H2DS13",
  "H2DS14"
)

# Scale developed by Bruefach & Reynolds https://www.sciencedirect.com/science/article/pii/S0049089X21001447?via%3Dihub#fn1
add_perceived_isolation_scale <- c(
  ### From Bruefach & Reynolds' Scale
  "S62E",    #I feel like I am part of this school
  "H1ED19", #I feel close to people at this school
  "S62I", #I am happy to be at this school
  "H1FS17", #You felt that people disliked you
  "H1FS14", # People were unfriendly to you
  "H1FS13", # You felt lonely
  ### Misc variables
  "H1DA7" # During the past week, how many times did you just hang out with friends?
)

```

```{r}

```


> ADD HEALTH Notes: School-level teacher & classroom race data are behind a paywall

## NLSY

```{r}
nlsy_diploma <- c(
  "HGC_DLI", # Highest grade completed as of date of last interview.
  "HSTDEGREE2018"  # Highest academic degree received as of the 2018 interview.
)

nlsy_school <- c(
  # It's easy to make friends at this school.
  "CS884143", 
  "CS906463", 
  "CS926611", 
  "CS942243", 
  "CS960613", 
  "CS98023A",	
  "CSAS023A",
  # Most of the teachers are willing to help with personal problems.
  "CS884144", 
  "CS906465", 
  "CS926613", 
  "CS942245", 
  "CS960615", 
  "CS98023B",	
  "CSAS023B",
  # Most of my classes are boring.
  "CS884145", 
  "CS906467", 
  "CS926615", 
  "CS942247", 
  "CS960617", 
  "CS98023C",	
  "CSAS023C",
  # I don't feel safe at this school.
  "CS884146", 
  "CS906469", 
  "CS926617", 
  "CS942249", 
  "CS960619", 
  "CS98023D",	
  "CSAS023D",
  # Most of my teachers don't know their subjects well.
  "CS884147", 
  "CS906471", 
  "CS926619", 
  "CS942251", 
  "CS960621", 
  "CS98023E",	
  "CSAS023E",
  # You can get away with almost anything at this school.
  "CS884148", 
  "CS906473", 
  "CS92662", 
  "CS942253", 
  "CS960623", 
  "CS98023F",	
  "CSAS023F",
  # My schoolwork requires me to think to the best of my abilities.
  "CS884149", 
  "CS906475", 
  "CS926623", 
  "CS942255", 
  "CS960625", 
  "CS98023G",	
  "CSAS023G",
  # At this school, a person has the freedom to learn what interests him or her.
  "CS884150", 
  "CS906477", 
  "CS926625", 
  "CS942257", 
  "CS960627", 
  "CS98023H",	
  "CSAS023H"
)


```



# Creating Long Dataframes

```{r Creating Long Dataframes}
years_names <- 1:14 
for (i in 1:14) {
  years_names[i] <- paste0("ever_", i*2+1986)
}

nlsy_longitudinal <- nlsy %>% 
  pivot_longer(cols = all_of(years_names), names_to = "years")
nlsy_longitudinal$value <- tolower(nlsy_longitudinal$value) %>% 
  factor()
nlsy_longitudinal$years <- sub("ever_","",  nlsy_longitudinal$years) %>% 
  as.numeric()
nlsy_longitudinal <- nlsy_longitudinal %>% 
  mutate(
    school_age = years - birth_year
  )
nlsy_longitudinal <- nlsy_longitudinal %>% 
  filter(school_age <= 18)
nlsy_l <- completeFun(nlsy_longitudinal, "value")
nlsy_l$school_group <- factor(nlsy_l$school_age)
nlsy_l$school_group <- forcats::fct_collapse(nlsy_l$school_group,
                                             elementary = c("3",  "4",  "5",  "6" , 
                                                            "7",  "8" , "9" , "10"),
                                             middle = c("11", "12" ,"13" ), 
                                             high = c("14","18"))
table(nlsy_l$school_group)
```

# Exporting Data

```{r}
save(nlsy, add, nlsy_l, file= "data.rdata")
save(NLSY_MINI_IPIP_QUESTIONS, NLSY_PERSONALITY_COR, NLSY_PERSONALITY_CORMAT_FULL, ADD_MINI_IPIP_COR, add_mini_ipip_questions, file = "personality_measures.rdata")
```
